{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}

### 2016-09-28 ident_manu.Rnw
added ident all figure, L1 to heurist tables

# format tocal_all for identifiability table
totab <- lapply(tocal_all, function(x) x$top) %>% 
  reshape2::melt(., id.var = names(.[[1]])) %>%
  mutate(Parameter = as.character(Parameter)) %>% 
  mutate(
    stts = factor(L1, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = cats_lev, labels = cats_lab), 
    Parameter = par_txt(Parameter, frm = 'tex')
  ) %>% 
  filter(coll < 15) %>% 
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank, '\\textsubscript{', Category, '}')
    ) %>%
  select(stts, Selected, Parameter, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\scriptsize{', Parameter, '}')
    ) %>% 
  arrange(stts, Selected)

# final table formatting
stts <- as.character(totab$stts)
sels <- totab$Selected
cap_val<- 'Parameter identifiability (as $\\gamma$, \\cref{gameq}) for relevant state variables.  Selections followed the second heuristic where parameters were selected independent of category from most to least sensitive until $\\gamma > 15$.  Rank describes the relative parameter sensitivity in each category for each state variable (O: optics, OM: organic matter, P: phytoplankton, T: temperature, Z: zooplankton). See \\cref{fig:heurist_stts} for a graphical illustration.'
colnms <- c('Parameter', 'Rank', '$\\gamma$')

@@

# format tocal_all for identifiability table
totab <- lapply(tocal_all, function(x) x$cattop) %>% 
  reshape2::melt(., id.var = names(.[[1]])) %>%
  mutate(Parameter = as.character(Parameter)) %>% 
  mutate(
    stts = factor(L1, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = cats_lev, labels = cats_lab), 
    Parameter = par_txt(Parameter, frm = 'tex')
  ) %>% 
  filter(coll < 15) %>% 
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank, '\\textsubscript{', Category, '}')
    ) %>%
  select(stts, Selected, Parameter, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\scriptsize{', Parameter, '}')
    ) %>% 
  arrange(stts, Selected)

# final table formatting
stts <- as.character(totab$stts)
sels <- totab$Selected
cap_val<- 'Parameter identifiability (as $\\gamma$, \\cref{gameq}) for relevant state variables.  Selections followed the third heuristic where parameters were selected equally within each category from most to least sensitive until $\\gamma > 15$.  Rank describes the relative parameter sensitivity in each category for each state variable (O: optics, OM: organic matter, P: phytoplankton, T: temperature, Z: zooplankton).'
colnms <- c('Parameter', 'Rank', '$\\gamma$')

@@

% identifiability boxplots
<<identbox, eval = T, fig.cap = 'Identifiability (as $\\gamma$, \\cref{gameq}) of parameter subsets for \\ac{do}.  Plots in (a) show identifiability by parameter categories and (b) shows identifiability by selecting the top 1 through 4 parameters in all categories.  Lines represent identifiability ranges for the possible combinations given the number of parameters in the set.  The temperature category is not shown because \\ac{do} was sensitive to only one parameter (i.e., $\\gamma = 1$).', fig.height = 7, fig.width = 8, echo = F, cache = TRUE>>=

@@

>>>>

# format tocal_all for identifiability table
totab <- lapply(tocal_all, function(x) x$top) %>% 
  reshape2::melt(., id.var = names(.[[1]])) %>%
  mutate(Parameter = as.character(Parameter)) %>% 
  mutate(
    stts = factor(L1, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = cats_lev, labels = cats_lab), 
    Parameter = par_txt(Parameter, frm = 'tex')
  ) %>% 
  filter(coll < 15) %>% 
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank, '\\textsubscript{', Category, '}')
    ) %>%
  select(stts, Selected, Parameter, error, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\scriptsize{', Parameter, '}')
    ) %>% 
  arrange(stts, Selected) %>% 
  rename(L1 = error) %>% 
  mutate(L1 = sapply(L1, scinot))

# final table formatting
stts <- as.character(totab$stts)
sels <- totab$Selected
cap_val<- 'Parameter identifiability (as $\\gamma$, \\cref{gameq}) for relevant state variables.  Selections followed the second heuristic where parameters were selected independent of category from most to least sensitive (L1, \\cref{l1}), until $\\gamma > 15$.  Rank describes the relative parameter sensitivity in each category for each state variable (O: optics, OM: organic matter, P: phytoplankton, T: temperature, Z: zooplankton). See \\cref{fig:heurist_stts} for a graphical illustration.'
colnms <- c('Parameter', 'L1', 'Rank', '$\\gamma$')

@@

# format tocal_all for identifiability table
totab <- lapply(tocal_all, function(x) x$cattop) %>% 
  reshape2::melt(., id.var = names(.[[1]])) %>%
  mutate(Parameter = as.character(Parameter)) %>% 
  mutate(
    stts = factor(L1, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = cats_lev, labels = cats_lab), 
    Parameter = par_txt(Parameter, frm = 'tex')
  ) %>% 
  filter(coll < 15) %>% 
  mutate(
    coll = sapply(coll, scinot), 
    Rank = sapply(rnk_incat, scinot), 
    Rank = paste0(Rank, '\\textsubscript{', Category, '}')
    ) %>%
  select(stts, Selected, Parameter, error, Rank, coll) %>%
  mutate(
    stts = as.character(stts),
    Parameter = paste0('\\scriptsize{', Parameter, '}')
    ) %>% 
  arrange(stts, Selected) %>% 
  rename(L1 = error) %>% 
  mutate(L1 = sapply(L1, scinot))

# final table formatting
stts <- as.character(totab$stts)
sels <- totab$Selected
cap_val<- 'Parameter identifiability (as $\\gamma$, \\cref{gameq}) for relevant state variables.  Selections followed the third heuristic where parameters were selected equally within each category from most to least sensitive (L1, \\cref{l1}), until $\\gamma > 15$.  Rank describes the relative parameter sensitivity in each category for each state variable (O: optics, OM: organic matter, P: phytoplankton, T: temperature, Z: zooplankton).'
colnms <- c('Parameter', 'L1', 'Rank', '$\\gamma$')

@@

% identifiability plot
<<identplo, eval = T, fig.cap = 'Identifiability (as $\\gamma$, \\cref{gameq}) of parameter subsets for \\ac{do}.  Plots in (a) show identifiability by parameter categories and (b) shows identifiability by selecting the top 1 through 4 parameters in all categories.  Lines represent identifiability ranges for the possible combinations given the number of parameters in the set.  The temperature category is not shown because \\ac{do} was sensitive to only one parameter (i.e., $\\gamma = 1$).', fig.height = 7, fig.width = 8, echo = F, cache = TRUE>>=

@@

% identifiability plot, all state variables
<<identploall, eval = T, fig.cap = 'Average identifiability (as $\\gamma$, \\cref{gameq}) of parameter subsets for all state variables.  Plots in (a) show identifiability by parameter categories and (b) shows identifiability by selecting the top 1 through 4 parameters in all categories.  Identifiability was averaged for all combinations in a parameter set to evaluate relative differenes between state variables.  The temperature category is not shown because all state variables were sensitive to only one parameter (i.e., $\\gamma = 1$, see \\cref{fig:inistreval}).', fig.height = 8, fig.width = 9, echo = F, cache = TRUE>>=

# state variable, categories, and tops to plot (no temp ident because only one parameter)
catnms <- c('Optics', 'Organic Matter', 'Phytoplankton', 'Zooplankton')
toplevs <- paste0('top', 1:4)
toplabs <- paste('Top', 1:4)

stts_lev <- c('NH4', 'Chla_mg_tot', 'O2', 'irradiance', 'NO3', 'OM1', 'OM2', 'PO4')
stts_lab <- c('Ammonium', 'Chlorophyll', 'Dissolved Oxygen', 'Irradiance', 'Nitrate', 'OM1', 'OM2', 'Phosphate')

toplo <- sapply(1:length(cat_ident_all), function(x){

  out <- cat_ident_all[[x]] %>% 
    .[names(.) %in% catnms] %>% 
    lapply(function(y) dplyr::select(y, N, coll)) %>% 
    reshape2::melt(., id.var = c('N', 'coll')) %>% 
    rename(Category = L1) %>% 
    mutate(Category = factor(Category, 
      levels = catnms,
      labels = catnms
      )) %>% 
    filter(N <= 25) %>% # reduced for plot space
    group_by(N, Category) %>% 
    summarise(
      ave = mean(coll),   
      med = median(coll),
      min = min(coll), 
      max = max(coll)
      )
  data.frame(stvar = names(cat_ident_all)[x], out, stringsAsFactors = F)

  }, simplify = F) %>% 
  do.call('rbind', .) %>% 
  mutate(`State variable` = factor(stvar, levels = stts_lev, labels = stts_lab))

stvars <- unique(toplo$`State variable`)
mycols <- RColorBrewer::brewer.pal(8, 'Spectral')
names(mycols) <- stvars
fils <- scale_fill_manual(name = 'State variable', values = mycols)

lncol <- 'darkgrey'
sz <- 3

p1 <- ggplot(toplo, aes(x = N, group = N)) + 
  geom_line(aes(y = ave, group = `State variable`), colour = lncol) + 
  geom_point(aes(y = ave, fill = `State variable`), shape = 21, colour = lncol, 
    size = sz
    ) + 
  fils + 
  facet_wrap(~ Category, ncol = 1, scales = 'free_y') +
  mytheme +
  scale_y_continuous(trans = 'log10') + 
  theme(legend.position = 'none') + 
  ggtitle('(a) By parameter category')

toplo2 <- sapply(1:length(top_ident_all), function(x){

  out <- top_ident_all[[x]] %>% 
    .[names(.) %in% toplevs] %>% 
    reshape2::melt(id.var = c('N', 'coll')) %>% 
    select(N, coll, L1) %>% 
    mutate(L1 = factor(L1, levels = toplevs, labels = toplabs)) %>% 
    rename(npercat = L1) %>% 
    group_by(N, npercat) %>% 
    summarise(
      ave = mean(coll), 
      med = median(coll),
      min = min(coll), 
      max = max(coll)
    )
  data.frame(stvar = names(top_ident_all)[x], out, stringsAsFactors = F)

  }, simplify = F) %>% 
  do.call('rbind', .) %>% 
  mutate(`State variable` = factor(stvar, levels = stts_lev, labels = stts_lab))

p2 <- ggplot(toplo2, aes(x = N, group = N)) + 
  geom_line(aes(y = ave, group = `State variable`), colour = lncol) + 
  geom_point(aes(y = ave, fill = `State variable`), shape = 21, colour = lncol, 
    size = sz
    ) + 
  fils + 
  facet_wrap(~ npercat, ncol = 1, scales = 'free_y') +
  mytheme +
  scale_y_continuous(trans = 'log10') + 
  ggtitle('(b) By top parameters, all categories')

grid.arrange(
  arrangeGrob(p1, p2, ncol = 2, bottom = 'Parameters in set'),
  left = 'Identifiability',
  pleg, ncol = 2, widths = c(1, 0.2)
)
@
